\section{Introduction}

\subsection{Background: the 5G core has become software}

For decades, a mobile operator's core network was built from specialised,
purpose-built hardware: a particular box routed subscriber traffic, another
managed mobility, another enforced policy. Each was a closed appliance from a
single vendor. With \textbf{5G Standalone (SA)}, this has changed completely.
The 3GPP 5G core \cite{ts23501} is defined as a set of \textbf{Network
Functions (NFs)} --- the User Plane Function (UPF), the Access and Mobility
Management Function (AMF), the Session Management Function (SMF), the Network
Repository Function (NRF), and others --- that communicate over open,
service-based interfaces. Crucially, these NFs are now \emph{software}. The
telecom industry packages them as containers and runs them on Kubernetes,
turning them into \textbf{Cloud-Native Network Functions (CNFs)}.

This shift is the cloud-native culmination of \textbf{Network Functions
Virtualisation (NFV)}, the architecture standardised by ETSI \cite{etsi002}
that decouples network functions from the hardware they used to run on. The
hardware and virtualisation layer underneath the NFs is called the
\textbf{NFV Infrastructure (NFVI)}; the system that manages it is the
\textbf{Virtual Infrastructure Manager (VIM)}. In the cloud-native era, the
NFVI is a pool of servers and the VIM is Kubernetes.

\subsection{Problem statement}

A 5G core cannot simply be dropped onto an ordinary Kubernetes cluster and be
expected to work. The NFs impose demands that a default cluster does not
satisfy:

\begin{itemize}
    \item The \textbf{UPF} forwards subscriber data and therefore needs its
          own \emph{user-plane} network interfaces (3GPP's N3 towards the
          radio side and N6 towards the Internet), kept separate from ordinary
          cluster traffic. A standard pod has only one interface --- not
          enough.
    \item The stateless control functions must \textbf{scale} with signalling
          load, and the whole core must be \textbf{observable} so an operator
          can see whether it is healthy.
    \item Subscriber state must live somewhere, and the different traffic
          classes (management, control signalling, user data) must not be
          mixed on one network.
\end{itemize}

Building and operating a production NFVI that meets these requirements is
expensive and complex --- well beyond what a student can stand up to learn
from. This project therefore asks a focused question:
\emph{can a faithful, miniature NFVI platform --- one that genuinely satisfies
the requirements of the 5G core --- be built on a single laptop, and can a
real 5G core actually run on it?}

The deliverable is explicitly an \textbf{infrastructure platform} (an NFVI /
container-as-a-service layer for CNFs), \emph{not} a new network application.
The 5G core is the demanding tenant we use to prove the platform is real.

\subsection{Objectives}

\begin{enumerate}
    \item Provision a multi-node Kubernetes (k3s) cluster entirely through
          Infrastructure-as-Code (Vagrant + Ansible), acting as the NFVI and
          VIM.
    \item Provide the \textbf{multi-NIC, multi-plane} networking the 5G
          user plane requires, using a secondary CNI
          (Multus + macvlan + Whereabouts), with management, cluster, and
          user-plane traffic strictly separated.
    \item Deploy a \textbf{real 5G SA core} (Open5GS) as a set of CNFs, and
          simulate the radio access network and a subscriber with UERANSIM.
    \item Demonstrate, end to end, a UE \textbf{registering} with the core and
          establishing a \textbf{PDU session} that carries real user-plane
          traffic through the UPF.
    \item Provide elastic \textbf{auto-scaling} of stateless NFs and full
          \textbf{observability} (Prometheus + Grafana) of the running core,
          all within a hard 16~GB RAM budget.
\end{enumerate}

\subsection{Report organisation}

Section~2 explains, for a reader new to the topic, the ETSI NFV framework and
the 5G core as CNFs. Section~3 derives, from the 5G core's needs, the precise
requirements the platform must meet, and maps each onto a Kubernetes mechanism.
Section~4 presents the system architecture --- including where each component
runs and how the design copes with failure. Section~5 details the
implementation, culminating in the deployment of Open5GS. Section~6 evaluates
the platform through five test cases driven by the live 5G core. Section~7
concludes and discusses limitations and future work.
