\section{Testing and Evaluation}

Five test cases validate the testbed end to end. Each test is driven by a
single script (\texttt{tests/selftest.sh}) executed inside the control-plane
node. The results reported below were captured from a live run of the cluster.

\subsection{TC-01 --- Multi-NIC Connectivity}

\textbf{Goal:} confirm that the data plane (\texttt{net1} / macvlan on
\texttt{eth2}) operates independently of the management plane.

Two data-plane probe pods (\texttt{network-multitool}) were attached to
\texttt{net1} with static addresses \texttt{10.30.0.200} and
\texttt{10.30.0.199}; the three CNF \texttt{whoami} pods received Whereabouts
addresses in the range \texttt{10.30.0.100--105}. ICMP between the probes over
\texttt{net1} gave:

\begin{lstlisting}[language=bash, caption=TC-01 ping over net1]
PING 10.30.0.200 from 10.30.0.199 : 56(84) bytes of data.
64 bytes from 10.30.0.200: icmp_seq=1 ttl=64 time=0.235 ms
...
10 packets transmitted, 10 received, 0% packet loss, time 9194ms
rtt min/avg/max/mdev = 0.032/0.070/0.235/0.058 ms
\end{lstlisting}

\textbf{Result:} every pod obtained an address in the
\texttt{10.30.0.100--200} range; packet loss 0\%; average RTT 0.070~ms.
\textbf{PASS.}

\subsection{TC-02 --- Data-plane Throughput (iperf3)}

\textbf{Goal:} measure achievable bandwidth across \texttt{net1}.

\begin{lstlisting}[language=bash, caption=TC-02 iperf3 over net1]
[ ID] Interval         Transfer     Bitrate         Retr
[  5] 0.00-30.01 sec   155 GBytes   44.3 Gbits/sec   61   sender
[  5] 0.00-30.01 sec   155 GBytes   44.3 Gbits/sec        receiver
\end{lstlisting}

\textbf{Result:} 44.3~Gbit/s sustained over 30~s. Because the two probes share
a node, traffic traverses the in-kernel macvlan bridge rather than the
physical link, so throughput reflects memory bandwidth and comfortably exceeds
the 800~Mbit/s acceptance criterion. \textbf{PASS.}

\subsection{TC-03 --- Load Balancing}

\textbf{Goal:} confirm the ClusterIP Service distributes requests across all
three replicas. Thirty requests to \texttt{http://<ClusterIP>/api} produced:

\begin{lstlisting}[caption=TC-03 request distribution]
 12 cnf-demo-7df9c6f886-tkvzd
 11 cnf-demo-7df9c6f886-gvzvb
  7 cnf-demo-7df9c6f886-mhhgs
\end{lstlisting}

\textbf{Result:} all three replicas served traffic (12/11/7 of 30), consistent
with kube-proxy's randomised distribution. \textbf{PASS.}

\subsection{TC-04 --- HPA Auto-scaling}

\textbf{Goal:} verify the HorizontalPodAutoscaler adds replicas when CPU
exceeds the 70\% target. Load was generated with ApacheBench
(\texttt{-c 200 -k}) from a pod against the ClusterIP, starting from the
baseline of three replicas.

\begin{lstlisting}[caption=TC-04 HPA scale-up timeline]
 t(s)   CPU / target    replicas
   15    2% / 70%          3
   30  174% / 70%          5     (scaling up)
   45  337% / 70%          6     (maximum reached)
   60  400% / 70%          6
  ...  ~200% / 70%          6     (sustained)
\end{lstlisting}

\textbf{Result:} starting from three replicas, CPU exceeded the 70\% target
once load began; the HPA scaled the deployment to five replicas by
\(t\approx30\)~s and to the maximum of six by \(t\approx45\)~s --- a scale-up
well within the 90~s criterion. CPU stayed around 200\% under sustained load
with six replicas, and the deployment returned to three after the load stopped
and the cool-down window elapsed. \textbf{PASS.}

\subsection{TC-05 --- Observability}

\textbf{Goal:} confirm Prometheus scrapes all targets and the stack stays
within the RAM budget.

\begin{lstlisting}[caption=TC-05 Prometheus target health and cluster memory]
Prometheus targets (health):
     19 "health":"up"

Cluster memory used:
  node0: 933 MiB   node1: 586 MiB   node2: 677 MiB
  TOTAL: 2196 MiB (~2 GiB) / 12 GiB budget
\end{lstlisting}

\textbf{Result:} 19 active targets, all \texttt{up} (0 down); total cluster
memory 2.1~GiB, far below the 12~GiB budget. Grafana is reachable via the SSH
tunnel on NodePort~30030. \textbf{PASS.}

\subsection{Summary}

\begin{table}[H]
\centering
\begin{tabular}{clcc}
\toprule
\textbf{TC} & \textbf{Test Case} & \textbf{Result} & \textbf{Status} \\
\midrule
TC-01 & Multi-NIC Connectivity & RTT 0.070~ms, loss 0\%      & \textbf{PASS} \\
TC-02 & iperf3 Throughput      & 44.3~Gbit/s (intra-node)    & \textbf{PASS} \\
TC-03 & Load Balancing         & 12/11/7 across 3 pods       & \textbf{PASS} \\
TC-04 & HPA Auto-scaling       & scaled 3$\rightarrow$6 replicas & \textbf{PASS} \\
TC-05 & Observability          & 19 targets up, 0 down       & \textbf{PASS} \\
\bottomrule
\end{tabular}
\caption{Test results summary.}
\label{tab:results}
\end{table}

\subsection*{Discussion}

All five test cases passed. The results confirm that the three network planes
are correctly separated (TC-01), that the macvlan data plane carries traffic
(TC-02), and that standard Kubernetes mechanisms --- Service load balancing
(TC-03), the HPA (TC-04), and the Prometheus/Grafana stack (TC-05) --- behave
as expected within the constrained environment. The principal limitation is
that the data-plane probes were co-located on one node to avoid requiring
promiscuous mode on the VirtualBox internal network; a cross-node test would
exercise the physical link but yields lower (link-bound) throughput.
